---
title: "Final Project"
author: "Celeste Kim"
date: "9/2/2019"
output:
  html_document: default
  pdf_document: default
---
### Data Science 250 Midterm Presentation

```{r echo=FALSE}
library(tidyverse)
library(dplyr)
rawdata <- read.csv("Collisions.csv")
df <- as_tibble(rawdata)
```

### Data

- https://catalog.data.gov/dataset/collisions
- Data set regarding all collisions from 2004 - present in the city of Seattle

### Initial Questions

- To what degree do driving conditions(road condition, light condition, weather) affect the chances of collisions?
- Do some parts of the road(intersection, block, alley) have more tendency to cause collisions?
- How has different variables of collisions changed over the time period of the data set?

### New/Altered Questions

## - How are driving conditions(road condition, light condition, weather) associated with the severity of collisions?

```{r echo=FALSE}
#selects road, light conditions, weather, and the severity of the collision
environment <- df %>% 
  select(ROADCOND, LIGHTCOND, WEATHER, SEVERITYCODE)

#creates roadcond  
  roadcond <- environment %>%
    select(ROADCOND, SEVERITYCODE) %>%
    filter(ROADCOND == "Dry"|
            ROADCOND == "Ice"|
            ROADCOND == "Oil"|
            ROADCOND == "Sand/Mud/Dirt"|
            ROADCOND == "Snow/Slush"|
            ROADCOND == "Standing Water"|
            ROADCOND == "Wet"
           ) %>%
    filter(!(SEVERITYCODE == 0))
  
  roadcond
#creates countplot for roadcond
    ggplot(roadcond, mapping = aes(x = ROADCOND, y = SEVERITYCODE)) + geom_count() + scale_size_area( max_size = 20)
```



```{r echo=FALSE}
#creates lightcond
    lightcond <- environment %>%
      select(LIGHTCOND, SEVERITYCODE) %>%
      filter(LIGHTCOND == "Dark - No Street Lights" |
               LIGHTCOND == "Dark - Street Lights Off" |
               LIGHTCOND == "Dark - Street Lights On" |
               LIGHTCOND == "Dawn" |
               LIGHTCOND == "Daylight" |
               LIGHTCOND == "Dusk" 
               ) %>%
      filter(!(SEVERITYCODE == 0))
  lightcond
#creates countplot for lightcond
    ggplot(lightcond, mapping = aes(x = LIGHTCOND, y = SEVERITYCODE)) + geom_count() + scale_size_area( max_size = 20 )
```

Light Conditions(In Order)

- Dark - No Street Lights

- Dark - Street Lights Off

- Dark - Stree Lights On

- Dawn

- Daylight

- Dusk

```{r echo=FALSE}
#creates weather
  weather <- environment %>%
    select(WEATHER, SEVERITYCODE) %>%
    filter(WEATHER == "Blowing Sand or Dirt or Snow" |
             WEATHER == "Clear or Partly Cloudy"|
             WEATHER == "Fog/Smog/Smoke" |
             WEATHER == "Overcast" |
             WEATHER == "Raining" |
             WEATHER == "Severe Crosswind" |
             WEATHER == "Sleet/Hail/Freezing Rain" |
             WEATHER == "Snowing"
           ) %>%
    filter(!(SEVERITYCODE == 0))
  
  weather
#creates countplot for weather
    ggplot(weather, mapping = aes(x = WEATHER, y = SEVERITYCODE)) + geom_count() + scale_size_area( max_size = 20)
```

Weathers(In Order)

- Blowing Sand or Dirt Snow

- Clear or Partly Cloudy

- Fog/Smog/Smoke

- Overcast

- Raining

- Severe Crosswind

- Sleet/Hail/Freezing Rain

- Snowing


## - How is the the type of the road associated with the severity of collisions? (Trevor’s Suggestion)

```{r echo=FALSE}
#selects type of road and compares it to severity (fatality, serious injury, injury, property damage)
road_type_severity <- df %>%
  select(ADDRTYPE, SEVERITYCODE) %>%
  filter(!(SEVERITYCODE == 0)) %>%
  group_by(ADDRTYPE) %>%
  summarise(fatality_rate = mean(SEVERITYCODE == "3") * 100,
            serious_injury_rate = mean(SEVERITYCODE == "2b") * 100,
            injury_rate = mean(SEVERITYCODE == 2 ) * 100,
            prop_dmg_rate = mean(SEVERITYCODE == 1) * 100) %>%
  filter(ADDRTYPE == "Block" |
        ADDRTYPE == "Alley" |
        ADDRTYPE == "Intersection" )
road_type_severity

ggplot(road_type_severity, aes(x = ADDRTYPE, y = fatality_rate)) + geom_col()
ggplot(road_type_severity, aes(x = ADDRTYPE, y = serious_injury_rate)) + geom_col()
ggplot(road_type_severity, aes(x = ADDRTYPE, y = injury_rate)) + geom_col()
ggplot(road_type_severity, aes(x = ADDRTYPE, y = prop_dmg_rate)) + geom_col()
```

## - How has different variables of collisions changed over the time period of the data set? (Not Changed)

```{r echo=FALSE}
#separates date into year, month, and date and compares it to severity
date_severity <-  df %>%
  select(INCDATE, SEVERITYCODE) %>%
  separate(INCDATE, into = c("year", "month", "date"), sep = "-", convert = TRUE) %>%
  separate(date, into = c("date", "idk"), sep = "T", convert = TRUE) %>%
  select(-idk) %>%
  arrange(year, month, date)
date_severity

date_severity2 <- date_severity %>%
  mutate(time = year + month / 13 + date / 310) %>%
  select(time, SEVERITYCODE) 
  
ggplot(date_severity2, mapping = aes(x = time, y = SEVERITYCODE)) + geom_point(position = "jitter")
```

## - The relationship between the locations in the city of Seattle and collisions (Jeong’s Suggestion)

## Difficulties

- Thankfully, most of the data set was tidy and did not require a lot of data wrangling
  * Examples of difficulties, however, included variable called INCDATE, which lumped together year, month, date, and an unrecognizable string together, which I had to use the "separate()" function to tidy
- Still having difficulties finding a method to create a visual incorporating the map of Seattle and markers indicating the frequency of collisions
- A lot of Unknown/implicitly missing values that require a lot of filtering
- Too much data (e.g. time vs severity)

## Looking Forward

- Continue research with the visual incorporating the map of Seattle and markers indicating the frequency of collisions
- Analyze data to potentially find causation, not just association
- Compare more variables with time
